---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# distops

<!-- badges: start -->
[![R-CMD-check](https://github.com/LMJL-Alea/distops/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/LMJL-Alea/distops/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/LMJL-Alea/distops/branch/master/graph/badge.svg)](https://app.codecov.io/gh/LMJL-Alea/distops?branch=master)
[![pkgdown](https://github.com/LMJL-Alea/distops/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/LMJL-Alea/distops/actions/workflows/pkgdown.yaml)
[![test-coverage](https://github.com/LMJL-Alea/distops/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/LMJL-Alea/distops/actions/workflows/test-coverage.yaml)
<!-- badges: end -->

The goal of **distops** is to provide a set of functions to compute distances between observations in a sample and to perform operations on distance matrices.

## Installation

You can install the development version of distops from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("LMJL-Alea/distops")
```

## Example

Let us compute the Euclidean distance matrix for the `cars` dataset. First, we convert the dataset to a list of observations:

```{r}
cars_list <- apply(cars, 1, \(.x) .x, simplify = FALSE)
cars_list
```

Next, we compute the Euclidean distance matrix via [`distops::dist()`]:

```{r}
cars_dist <- distops::dist(cars_list)
cars_dist
```

We can subset this matrix using the `[` operator:

```{r}
cars_dist[2:3, 7:12]
```
