---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# distops

<!-- badges: start -->
[![R-CMD-check](https://github.com/LMJL-Alea/distops/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/LMJL-Alea/distops/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/LMJL-Alea/distops/branch/master/graph/badge.svg)](https://app.codecov.io/gh/LMJL-Alea/distops?branch=master)
[![pkgdown](https://github.com/LMJL-Alea/distops/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/LMJL-Alea/distops/actions/workflows/pkgdown.yaml)
[![test-coverage](https://github.com/LMJL-Alea/distops/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/LMJL-Alea/distops/actions/workflows/test-coverage.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/distops)](https://CRAN.R-project.org/package=distops)
<!-- badges: end -->

The goal of **distops** is to provide a set of functions to compute distances between observations in a sample and to perform operations on distance matrices.

## Installation

You can install the development version of distops from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("LMJL-Alea/distops")
```

## Features

```{r}
library(distops)
```

### Package developement

We provide two functions for package developers to help with defining efficient implementation of the `dist` functions for custom distances. Namely:

- `use_distops()` setups a package to use **distops** for computing distances. 
In particular, it creates a `src/` directory with a `Makevars` file and a
`Makevars.win` file. It also creates a `R/distops-package.R` file with the
appropriate **roxygen2** tags so that the `NAMESPACE` file is modified to add
the `importFrom()` directives for the Rcpp and RcppParallel packages and the
`useDynLib()` directive for packages with compiled code. It finally modifies the
`DESCRIPTION` file to add **Rcpp**, **RcppParallel** and **distops** to the
`Imports` and `LinkingTo` fields and GNU make to the `SystemRequirements` field.
- `use_distance()` creates R and C++ files for easy implementation of custom 
distances.

### Subset operator

Let us compute the Euclidean distance matrix for the `iris` dataset:

```{r}
D <- dist(iris[, 1:4], method = "euclidean")
```

We can subset this matrix using the `[` operator. We can either provide the same
indices for rows and columns in which case it return another object of class
`dist`:

```{r}
D[1:3, 1:3]
```

Or we can provide different indices for rows and columns in which case it
returns a dense matrix:

```{r}
D[2:3, 7:12]
```

The subsetting operation is fully parallelized using the **RcppParallel** 
package. It is also memory efficient as it does not copy the original distance 
matrix.

### Medoid computation

The medoid of a sample is the observation that minimizes the sum of distances to
all other observations. The `find_medoids()` function computes the medoid of a
sample for a given distance. It takes advantage of the **RcppParallel** package
to compute the medoid in parallel.

```{r}
find_medoids(D)
```

If the `memberships` argument is provided, it returns the medoid for each
cluster.

```{r}
find_medoids(D, memberships = as.factor(rep(1:3, each = 50L)))
```

## Future work

* Pass a list instead of a matrix to be more general?
* Use Arrow parquet format to store distance matrix in multiple files when 
sample size exceeds 10,000 or something like that.
* Use Arrow connection to read in large data.
* Add Progress bar.
